---
title: "Data Import"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


This document will show how to import data.



## Import the FAS Litters CSV

Import, and clean the variable names (make it all lower case) .
If use `library(janitor)` in advance, you can directly use `clean_names` as a function.

```{r}
litters_df = read_csv("data/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)
```


## Look at the dataset

```{r}
litters_df

head(litters_df)
tail(litters_df, 10)
```

```{r eval = FALSE}
view(litters_df)
```

## Import FAS Pups

Use relative paths.

```{r}
pups_df = read_csv("data/FAS_pups.csv")

pups_df = janitor::clean_names(pups_df)

pups_df
```

To use absolute path: `pups_df = read_csv("/Users/.../data/FAS_pups.csv")`
If move the files, this won't works anymore.


## Look at read_csv options

col_names and skipping rows.

Use `col_names = FALSE` to remove column names.
Use `skip = 1` to skip the first line (depends on the dataset).

```{r}
litters_df =
  read_csv(
    file = "data/FAS_litters.csv",
    skip = 1
  )
```

What about missing data.

Use `mean(x, na.rm = FALSE)` to remove NA when calculating mean.

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", ".")
  )

litters_df = janitor::clean_names(litters_df)

pull(litters_df, gd0_weight)
```


What if we code `group` as a factor variable?

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group = col_factor()
    )
  )
```


## Import an excel file

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet= "mlb11")
```

## Import a SAS file

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```


## Never use read.csv()

It will print everything, and not helpful with coding.
read_csv will give a tibble instead of data frame.

Never do this either:
```{r}
litters_df$Group
```
Don't take variable out of data frame.



